Speculation in foreign exchange: noise or information

### 2015-05-09 Knitr/Paper.tex
Updated the files. Now need to read through and make sure that it is ready to be sent off. What is the angle?

\documentclass[12pt, a4paper, oneside]{article} % Paper size, default font size and one-sided paper
%\usepackage[dcucite]{harvard}
\usepackage{tikz}
\usetikzlibrary{shapes, shadows, arrows}
\usepackage{rotating}
\usepackage{amsmath}
%\usepackage{setspace}
\usepackage{pdflscape}
\usepackage[flushleft]{threeparttable}
\usepackage{multirow}
\usepackage[comma, sort&compress]{natbib}% Use the natbib reference package - read up on this to edit the reference style; if you want text (e.g. Smith et al., 2012) for the in-text references (instead of numbers), remove 'numbers' 
\usepackage{graphicx}
\graphicspath{{./Figures/}} % Specifies the directory where pictures are stored
%\bibliographystyle{plainnat}
\usepackage{listings}
\bibliographystyle{agsm}
\usepackage[colorlinks = true, citecolor = blue, linkcolor = blue]{hyperref}
%\hypersetup{urlcolor=blue, colorlinks=true} % Colors hyperlinks in blue - change to black if annoying
\begin{document}
%\renewcommand[\harvardurl]{URL: \url}
\title{Speculation in foreign exchange: noise or information}
\author{Rob Hayward\footnote{University of Brighton Business School, Lewes Road, Brighton, BN2 4AT; Telephone 01273 642586.  rh49@brighton.ac.uk.  We are very grateful for the comments and suggestions made.  In particular, the EACES conference in Budapest between September 5th and 6th 2014, Andros Steinhoffer and.....}} 
\date{\today}
\maketitle
\begin{abstract}
An event study of extreme speculation is used to assess the Keynesian and Friedman views of speculation.  This is partly a test of the informational content of speculation.  Speculative extremes are ranked by weight of activity and intensity of sentiment.  In either case, the extreme says little about the future, suggesting that speculation is more than just noise and that there is some informational content that is being passed through to the price. A unique dataset of risk-reversal skew on option prices is used to measure the weight of speculative sentiment.  This is suplemented with regulatory data on the futures position of speculative traders.  
\end{abstract}

@@

This is a contrast between an idea of speculative activity as a nuisance that, at its most extreme, will likely to cause bubbles and reversal and will discourage fundamental long-term investors from participating in the market.  It is noise.   that does not emerge randomly, but feeds off itsef as a feedback of sentiment  encourages  trends and serial correlation in asset prices returns.  The alternative view regards speculation as an key lubricant of financial market as it provides liquidity and facilitates price discovery.

The aim of this paper is to understand mroe about the nature of speculation in the foreign exchange market by assess the relationship between the intensity or the weight of speculation and the movement of prices.  If speculators are uninformed and their activity can be regarded as noise, their most extreme sentiment and the time when speculators have the greatest weight in market activity should be those when price reversals are most likely to take place.  An event study is used to assess the relationship between speculative extremes and foreign exchange returns.  A unique set of option prices are used to measure the intensity of speculative sentiment.  US regulatory positions measure the prominence of speculators in the market.

\section{Literature}
From "How noise trading..." in the other folder. 
\citet{bloomfield2009noise} suggests that there is a difference between noise traders and liquidity traders.  Noise traders exert positive effects on liquidity but weaken price reversals. A tax on trading will reduce noise traders but will also reduce the influence of fundamental traders. Their experiment distinguishes between informed traders, noise traders and liquidity traders.  \emph{SOES bandits} are those taking advantage of the SOES system to trade quickly and take advantage of small price movements. There is a lot of focus on the effect that taxation may have on the different categories of trader. There is an experiment that allows much more to be held constant. There are three views of noise traders that are tested here:  noise traders are skill experts taking advantage of information in the order book or price movements;  rationally providing liquidity by acting as market-maker; acting irrationally as noise traders. Noise traders in this experiement appear to act as irrational contrarian traders. Noise traders refuce the efficiency of the market but increase liquidity.

Noise is related to rational expectations.  A mean zero, normally-distributed random variable influences supply citet{Muth1961Rational}.  Why do noise traders exist?  There are liquidity and behavioural explanations. There are also a number of models that show noise traders as momentum traders. The effect of noise traders on liquidity is unclear. They may supply liquidity by providing limit orders and making some returns for providing liquidity; they may increase bid-ask spreads by using market orders to take advantage of (potentially non-existant) informational advantages. Noise traders may increase (if they have some priviledged information, they add to the price pressure) or reduce efficiecy (if they increase liquiity and reduce the ratio of informed to uninformed traders).

Are day traders profitable? Are these the face of noise traders?  How does this show up in teh foreign exchange market?

\citet{da2010sum} Seek to measure invetor sentiment and assess how it affects securities prices. They create a index of investor sentiment from web searches and use this to predict return reversals.  The resutls show that sentiment can predict market turns. This is a contrast to traditional ways of measuring sentiment:  proxy for sentiment with market-based measures like volume, closed-end fund discount, IPO first-day returns, IPO volume, option implied volatities and mutual fund flows. Baker and Wurgaler (2007) have a comprehensive survey. This is more than just investors sentiment. There is an issue with the measurment of an input (sentiment) with an output (market price). Use of surveys like the University of Michigan consumer sentiment survey etc.  There is a negative contemporaneous correlation between FEARS and stock market returns. However, there is a reversal in the relationship after two days.  This is consistent with ``sentiment induced temporary mispricing'' \citet[p. 4]{da2010sum}.  There is a positive correlation between the FEARS index and bond prices.  However, thsi also reverses in following days. There is a relatioknship betwen the FEARS index and daily market volatility (realised volatlity and the VIX index). There is also evidence of inflows and outflows from mutual funds.  This may help to explain the contrast between evidence from the equity market and the bond market.

Literature.  Black (1986) highlights the positive effect of noise traders. Shliefer, Summers, DeLong etc identify the limits to arbitrage. The other writings are Battalio, Hatch and Jennings (1997), Harris and Schultz (1998) and Scheinkman and Xiong (2003).  Lowenstein and Willard (2006).  There is considerable debate about the precise role of noise trading in financial markets. There are two strands to the literature on noise trading that emerged in the 1980s:  microstricture literature will use the term "noise traders" and "liquidity traders" interchangeably to identify traders without information.  The motivation of these traders is not identified. Examples of this are Glosten and Milgrom (1985) and Kyle (1985). It is usually assumed to be hedging or liquidity. There is also a \emph{limits-to-arbitrage} literature that uses the term to identify traders that are not movivated by standard explanations of fundamental or liquidity provision. This paper seeks to distinguish between those that trade for liquidty or heging and those that trade because of psychological biases. Fisher Black ``People who trade on noise are willing to trade even though from an objective point of view they would be better off not trading. Perhaps they think the noise they are trading on is information. Or perhaps they just like to trade``.

In theory, traders may trade on mistaken fundamental information.  Shliefer and Summmer (1990) use momentum.  However, Kaniel, Saar and Titman (2006) find evidence of contrarian strategies. This is associated with short-term speculation. Does the activity drive the price away from fundamental value? Stiglitz (1990) is one of the arguments in favour of a Tobin tax.

There is a quesiton about whether experimental markets (like this paper) are superior to actual markets where the identification of trader types is difficult.  Theory - Delong, summers, shleifer etc,  Scheinkman and Xiong (2003).  Lowenstein and Willard (2006).  Empirical work:  (see, for example, Garvey and Murphy [2001], Linnainmaa [2003], Barber et. al. [2004], and Kaniel, Saar, and Titman [2006].

@@

Interesting difference between supplying and demanding liqudiity.  Limit orders supply liquidity, market orders demand liqudity.  The activity of liquidity traders will tend to change through the course of the session, moving from supplying to demanding liquidity. Noise traders tend to act in the opposite manner.  They tend to supply more liquidity as the session progresses. Black (1986) noise traders put noise into prices.  The effect of noise traders on liquidity: price impact of trades is lower when noise traders are present.  However, this is mostly in the temporary impact on prices rather than the permanent impact. Therefore, noise traders tend to reduce the amount of volatility in prices because market orders do not temporarily have an effect and then reverse.  The presence of noise traders will increase pricing errors. However, this depends on fundametal traders have valuable information.

There are a number of different ways that noise traders could behave:  activing as SOES bandits (using the NASDAQ small order execution system to make profits) or by using information about the order book; by providing liqudity like a market-maker as a service to liqudiity-traders as in Grossman Miller (1988);  as irrational noise traders, using trading rules or popular models (Shiller 1984, 1990) positive feedback or contrarian.

Footnote -  There is an empirical evidence that individual investors in various countries trade in a contrarian fashion (e.g., Choe, Kho, and Stulz [1999], Grinblatt and Keloharju [2000, 2001], Jackson [2003], Richards [2005], Kaniel, Saar, Titman [2006]). For experimental evidence consistent with contrarian behavior see 
Bloomfield, Tayler, and Zhou (2997)

Can noise traders be viewed as just random noise.  Dow and Gorton (2006) suggest that this is the case and that this is an outcrop of the rational expectations revolution. Informed traders profit by exploiting the diviations from equilibrium. This is the beneifts of Grossman-Stiglitz. What is the source of this noise?  What are the insitiutions? The importance of noise traders here is that they represent a zero-mean, normally distributed random vaiable.  This is also the description that is provided in Kyle (1985).  Liquidity traders trade due to imbalance in the timing of consumption or from portfolio considerations. Shleifer, Summers and DeLong have trading activity that is affected by sentiment, models or "fads".  This is important.  How the average may not be zero. There is a range of behavioural biases that may account for this (see Barberis and Thaler [2003]).The literaure on hetrogenous traders also focuses on noise traders as trend chasers.  These models (see for example Brock and Hommes [1998]; Lux [1995]; Lux and Marchesi [2000]) investigate bubbles and crashes, and feature fundamental traders versus noise traders. In these models, it is the noise traders who 
induce the aberrant market behavior.

\href{http://papers.ssrn.com.ezproxy.liv.ac.uk/sol3/papers.cfm?abstract_id=869827}{Do noise traders move markets?}  There is strong evidence of herding by individual investors. THere is also momentum. There is reversal:  stocks that were heavily bought one year underperform; stocks that were heavily sold, outperform. The opposite in the short run:  heavily bought one week generate large returns the next week.

\href{http://restud.oxfordjournals.org.ezproxy.liv.ac.uk/content/60/1/1.full.pdf+html}{Campbell and Kyle, Smart money, noise trading and stock market behaviour}. In response to the evidence that the volatilty of stock prices is too great to be a function of discounted expected future earnings (Shiller (1981),  LeRoy and Porter (1981), this paper assesses whether stocks are influnced by exogenous, serially correlated noise. In this model, the noise arises from the interaction of noise traders. 
% Does this now need something on the reason why the foreign exchange market is used?  Could be the Keynes speuclative activity, Fridman paper, liquidity etc.

@@

\section{Literature review}  
There has been an accumulation of evidence to suggest that financial asset prices do not evolve as a random walk.  Serial correlation or trends are evident suggesting that, contrary to the efficient market hypthesis, future price changes could be partially foreseen by the movement of past price changes.  There are a number of studies that find evidence that the reaction of stocks to information is less than instantaneous.  Ball and Brown seems to be the first study that identifies the upward drift in cumulative abnormal returns after some positive earnings news and downward drift in cumulative abnormal earnings after negative earnings news \citep{BallBrown}.  Bernard provides a summary of the evidence that stocks react slowly to earnings announcements.  He cites twenty examples of ``post earnings announcement drift'' \citep[p. 303]{BernardDrift}.  Part of this under-reaction comes from simply failing to identify what current earnings mean for subsequent quarters.    Bernard argues that the evidence is firmly on the side of earning reactions being too small and taking on average six months to be fully digested \citep[p. 305]{BernardDrift}.

@@

The idea that over-and-under shooting can be used to identify securities that have caught up in a wave of positive or negative sentiment has been repeated a number of times.  De Bondt and Thaler used the performance over the previous year to identify \emph{winners} and \emph{losers}, other have broken the portfolios into size-adjusted and beta-adjusted components to test the suggestion that the returns recorded are just a compensation for increased risk  \citep{Chopra1992}.  The main findings are that smaller firms account for the bulk of the gains but that overshooting is still evident.  There are other examples where a variety of measures, including the price-earnings ratio \citep{BasuPE} and cash flow to price \citep{Chan1991} have been used to assess the level of over-shooting.  It is shown that book-to-market and cash-flow ratios have the most positive impact on positive returns.  Lakonishok, Shleifer and Vishny provide an overview of \emph{contrarian} investment strategies \citep{LakonishokContrarian}.  This whole area is closely linked to the debate about the relative merits of \emph{value} and \emph{growth} stocks, with ample evidence from academic and practitioners that while growth stocks outperform value in the short-run, in line with the evidence of momentum, there are frequent reversals as over-reactions are reversed.

@@

\citep{LoMacKinlay1990} argue that one reason for the serial correlation in prices is the asynchronous trading system and the time lag between some stocks reacting to market news.  In other words, it could be that the underperforming stocks have not yet reacted to the news that the outperforming stocks have already incorporated.  However, this idea is rejected by Jagadeesh and Titman as a result of their method of decomposing returns into market and specific factors \citep[p. 73]{Jagadeesh}.  Indeed, Jagdeesh and Titman argue that the momentum that they find in stock price returns is consistent with delayed price reaction to firm-specific news \citep{Jagadeesh}

@@

With a high level of uncertainty, there is a certain scepticism about new information and a caution about changing an established opinion. The Psychologist Edwards describes his findings relative to the benchmark of Bayesian decision-making.  
\begin{quotation}
``An abundance of research has shown that human beings are conservative processors of fallible information.  Such experiments compare human behavior to the outputs of Bayes's theorem, the formally optimal rule about how opinions (that is, probabilities) should be revised on the basis of new information.  It turns out that opinion change is very \emph{orderly}, and usually proportional to number calculated from Bayes's theorem  - but it is insufficient in amount.  A convenient first approximation to the data would say that it takes anywhere from two to five observations to do one observation's worth of work in inducing a subject to change his opinions."  
\end{quotation}
\citep[p. 359]{KahnemanUncertainty}

>>>>

\documentclass[12pt, a4paper, oneside]{article} % Paper size, default font size and one-sided paper
%\usepackage[dcucite]{harvard}
\usepackage{tikz}
\usetikzlibrary{shapes, shadows, arrows}
\usepackage{rotating}
\usepackage{amsmath}
%\usepackage{setspace}
\usepackage{pdflscape}
\usepackage[flushleft]{threeparttable}
\usepackage{multirow}
\usepackage[comma, sort&compress]{natbib}% Use the natbib reference package - read up on this to edit the reference style; if you want text (e.g. Smith et al., 2012) for the in-text references (instead of numbers), remove 'numbers' 
\usepackage{graphicx}
\graphicspath{{./Figures/}} % Specifies the directory where pictures are stored
%\bibliographystyle{plainnat}
\usepackage{listings}
\bibliographystyle{agsm}
\usepackage[colorlinks = true, citecolor = blue, linkcolor = blue]{hyperref}
%\hypersetup{urlcolor=blue, colorlinks=true} % Colors hyperlinks in blue - change to black if annoying
\begin{document}
%\renewcommand[\harvardurl]{URL: \url}
\title{Speculation in foreign exchange: noise or information}
\author{Rob Hayward\footnote{University of Brighton Business School, Lewes Road, Brighton, BN2 4AT; Telephone 01273 642586.  rh49@brighton.ac.uk.  We are very grateful for the comments and suggestions made.  In particular, the EACES conference in Budapest between September 5th and 6th 2014, Andros Steinhoffer and.....}} 
\date{\today}
\maketitle
\begin{abstract}
An event study of extreme speculation is used to assess the Keynesian and Friedman views of speculation.  This is partly a test of the informational content of speculation.  Speculative extremes are ranked by weight of activity and intensity of sentiment.  In either case, the extreme says little about the future, suggesting that speculation is more than just noise and that there is some informational content that is being passed through to the price. A unique dataset of risk-reversal skew on option prices is used to measure the weight of speculative sentiment.  This is supplemented with regulatory data on the futures position of speculative traders.  
\end{abstract}

@@

This is a contrast between an idea of speculative activity as a nuisance that, at its most extreme, will likely to cause bubbles and reversal and will discourage fundamental long-term investors from participating in the market.  It is noise.   that does not emerge randomly, but feeds off itself as a feedback of sentiment  encourages  trends and serial correlation in asset prices returns.  The alternative view regards speculation as an key lubricant of financial market as it provides liquidity and facilitates price discovery.

The aim of this paper is to understand more about the nature of speculation in the foreign exchange market by assess the relationship between the intensity or the weight of speculation and the movement of prices.  If speculators are uninformed and their activity can be regarded as noise, their most extreme sentiment and the time when speculators have the greatest weight in market activity should be those when price reversals are most likely to take place.  An event study is used to assess the relationship between speculative extremes and foreign exchange returns.  A unique set of option prices are used to measure the intensity of speculative sentiment.  US regulatory positions measure the prominence of speculators in the market.

\section{Literature}
From "How noise trading..." in the other folder. 
\citet{bloomfield2009noise} suggests that there is a difference between noise traders and liquidity traders.  Noise traders exert positive effects on liquidity but weaken price reversals. A tax on trading will reduce noise traders but will also reduce the influence of fundamental traders. Their experiment distinguishes between informed traders, noise traders and liquidity traders.  \emph{SOES bandits} are those taking advantage of the SOES system to trade quickly and take advantage of small price movements. There is a lot of focus on the effect that taxation may have on the different categories of trader. There is an experiment that allows much more to be held constant. There are three views of noise traders that are tested here:  noise traders are skill experts taking advantage of information in the order book or price movements;  rationally providing liquidity by acting as market-maker; acting irrationally as noise traders. Noise traders in this experiment appear to act as irrational contrariety traders. Noise traders reduce the efficiency of the market but increase liquidity.

Noise is related to rational expectations.  A mean zero, normally-distributed random variable influences supply \citet{Muth1961Rational}.  Why do noise traders exist?  There are liquidity and behavioural explanations. There are also a number of models that show noise traders as momentum traders. The effect of noise traders on liquidity is unclear. They may supply liquidity by providing limit orders and making some returns for providing liquidity; they may increase bid-ask spreads by using market orders to take advantage of (potentially non-existent) informational advantages. Noise traders may increase (if they have some privileged information, they add to the price pressure) or reduce efficiency (if they increase liquidity and reduce the ratio of informed to uninformed traders).

Are day traders profitable? Are these the face of noise traders?  How does this show up in the foreign exchange market?

\citet{da2010sum} Seek to measure investor sentiment and assess how it affects securities prices. They create a index of investor sentiment from web searches and use this to predict return reversals.  The results show that sentiment can predict market turns. This is a contrast to traditional ways of measuring sentiment:  proxy for sentiment with market-based measures like volume, closed-end fund discount, IPO first-day returns, IPO volume, option implied volatilities and mutual fund flows. Baker and Wurgaler (2007) have a comprehensive survey. This is more than just investors sentiment. There is an issue with the measurement of an input (sentiment) with an output (market price). Use of surveys like the University of Michigan consumer sentiment survey etc.  There is a negative contemporaneous correlation between FEARS and stock market returns. However, there is a reversal in the relationship after two days.  This is consistent with ``sentiment induced temporary mispricing'' \citet[p. 4]{da2010sum}.  There is a positive correlation between the FEARS index and bond prices.  However, this also reverses in following days. There is a relationship between the FEARS index and daily market volatility (realised volatility and the VIX index). There is also evidence of inflows and outflows from mutual funds.  This may help to explain the contrast between evidence from the equity market and the bond market.

Literature.  Black (1986) highlights the positive effect of noise traders. Shliefer, Summers, DeLong etc identify the limits to arbitrage. The other writings are Battalio, Hatch and Jennings (1997), Harris and Schultz (1998) and Scheinkman and Xiong (2003).  Lowenstein and Willard (2006).  There is considerable debate about the precise role of noise trading in financial markets. There are two strands to the literature on noise trading that emerged in the 1980s:  microstricture literature will use the term "noise traders" and "liquidity traders" interchangeably to identify traders without information.  The motivation of these traders is not identified. Examples of this are Glosten and Milgrom (1985) and Kyle (1985). It is usually assumed to be hedging or liquidity. There is also a \emph{limits-to-arbitrage} literature that uses the term to identify traders that are not motivated by standard explanations of fundamental or liquidity provision. This paper seeks to distinguish between those that trade for liquidity or hedging and those that trade because of psychological biases. Fisher Black ``People who trade on noise are willing to trade even though from an objective point of view they would be better off not trading. Perhaps they think the noise they are trading on is information. Or perhaps they just like to trade``.

In theory, traders may trade on mistaken fundamental information.  Shliefer and Summmer (1990) use momentum.  However, Kaniel, Saar and Titman (2006) find evidence of contrariety strategies. This is associated with short-term speculation. Does the activity drive the price away from fundamental value? Stiglitz (1990) is one of the arguments in favour of a Tobin tax.

There is a question about whether experimental markets (like this paper) are superior to actual markets where the identification of trader types is difficult.  Theory - Delong, summers, shleifer etc,  Scheinkman and Xiong (2003).  Lowenstein and Willard (2006).  Empirical work:  (see, for example, Garvey and Murphy [2001], Linnainmaa [2003], Barber et. al. [2004], and Kaniel, Saar, and Titman [2006].

@@

Interesting difference between supplying and demanding liquidity.  Limit orders supply liquidity, market orders demand liquidity.  The activity of liquidity traders will tend to change through the course of the session, moving from supplying to demanding liquidity. Noise traders tend to act in the opposite manner.  They tend to supply more liquidity as the session progresses. Black (1986) noise traders put noise into prices.  The effect of noise traders on liquidity: price impact of trades is lower when noise traders are present.  However, this is mostly in the temporary impact on prices rather than the permanent impact. Therefore, noise traders tend to reduce the amount of volatility in prices because market orders do not temporarily have an effect and then reverse.  The presence of noise traders will increase pricing errors. However, this depends on fundamental traders have valuable information.

There are a number of different ways that noise traders could behave:  activating as SOES bandits (using the NASDAQ small order execution system to make profits) or by using information about the order book; by providing liquidity like a market-maker as a service to liquidity-traders as in Grossman Miller (1988);  as irrational noise traders, using trading rules or popular models (Shiller 1984, 1990) positive feedback or contrariety.

Footnote -  There is an empirical evidence that individual investors in various countries trade in a contrariety fashion (e.g., Choe, Kho, and Stulz [1999], Grinblatt and Keloharju [2000, 2001], Jackson [2003], Richards [2005], Kaniel, Saar, Titman [2006]). For experimental evidence consistent with contrariety behaviour see 
Bloomfield, Tayler, and Zhou (2997)

Can noise traders be viewed as just random noise.  Dow and Gorton (2006) suggest that this is the case and that this is an outcrop of the rational expectations revolution. Informed traders profit by exploiting the divinations from equilibrium. This is the benefits of Grossman-Stiglitz. What is the source of this noise?  What are the institutions? The importance of noise traders here is that they represent a zero-mean, normally distributed random viable.  This is also the description that is provided in Kyle (1985).  Liquidity traders trade due to imbalance in the timing of consumption or from portfolio considerations. Shleifer, Summers and DeLong have trading activity that is affected by sentiment, models or "fads".  This is important.  How the average may not be zero. There is a range of behavioural biases that may account for this (see Barberis and Thaler [2003]).The literature on heterogeneous traders also focuses on noise traders as trend chasers.  These models (see for example Brock and Hommes [1998]; Lux [1995]; Lux and Marchesi [2000]) investigate bubbles and crashes, and feature fundamental traders versus noise traders. In these models, it is the noise traders who 
induce the aberrant market behaviour.

\href{http://papers.ssrn.com.ezproxy.liv.ac.uk/sol3/papers.cfm?abstract_id=869827}{Do noise traders move markets?}  There is strong evidence of herding by individual investors. There is also momentum. There is reversal:  stocks that were heavily bought one year under perform; stocks that were heavily sold, outperform. The opposite in the short run:  heavily bought one week generate large returns the next week.

\href{http://restud.oxfordjournals.org.ezproxy.liv.ac.uk/content/60/1/1.full.pdf+html}{Campbell and Kyle, Smart money, noise trading and stock market behaviour}. In response to the evidence that the volatility of stock prices is too great to be a function of discounted expected future earnings (Shiller (1981),  LeRoy and Porter (1981), this paper assesses whether stocks are influenced by exogenous, serially correlated noise. In this model, the noise arises from the interaction of noise traders. 
% Does this now need something on the reason why the foreign exchange market is used?  Could be the Keynes speuclative activity, Fridman paper, liquidity etc.

@@

\section{Literature review}  
There has been an accumulation of evidence to suggest that financial asset prices do not evolve as a random walk.  Serial correlation or trends are evident suggesting that, contrary to the efficient market hypothesis, future price changes could be partially foreseen by the movement of past price changes.  There are a number of studies that find evidence that the reaction of stocks to information is less than instantaneous.  Ball and Brown seems to be the first study that identifies the upward drift in cumulative abnormal returns after some positive earnings news and downward drift in cumulative abnormal earnings after negative earnings news \citep{BallBrown}.  Bernard provides a summary of the evidence that stocks react slowly to earnings announcements.  He cites twenty examples of ``post earnings announcement drift'' \citep[p. 303]{BernardDrift}.  Part of this under-reaction comes from simply failing to identify what current earnings mean for subsequent quarters.    Bernard argues that the evidence is firmly on the side of earning reactions being too small and taking on average six months to be fully digested \citep[p. 305]{BernardDrift}.

@@

The idea that over-and-under shooting can be used to identify securities that have caught up in a wave of positive or negative sentiment has been repeated a number of times.  De Bondt and Thaler used the performance over the previous year to identify \emph{winners} and \emph{losers}, other have broken the portfolios into size-adjusted and beta-adjusted components to test the suggestion that the returns recorded are just a compensation for increased risk  \citep{Chopra1992}.  The main findings are that smaller firms account for the bulk of the gains but that overshooting is still evident.  There are other examples where a variety of measures, including the price-earnings ratio \citep{BasuPE} and cash flow to price \citep{Chan1991} have been used to assess the level of over-shooting.  It is shown that book-to-market and cash-flow ratios have the most positive impact on positive returns.  Lakonishok, Shleifer and Vishny provide an overview of \emph{contrariety} investment strategies \citep{LakonishokContrarian}.  This whole area is closely linked to the debate about the relative merits of \emph{value} and \emph{growth} stocks, with ample evidence from academic and practitioners that while growth stocks outperform value in the short-run, in line with the evidence of momentum, there are frequent reversals as over-reactions are reversed.

@@

\citep{LoMacKinlay1990} argue that one reason for the serial correlation in prices is the asynchronous trading system and the time lag between some stocks reacting to market news.  In other words, it could be that the under performing stocks have not yet reacted to the news that the outperforming stocks have already incorporated.  However, this idea is rejected by Jagadeesh and Titman as a result of their method of decomposing returns into market and specific factors \citep[p. 73]{Jagadeesh}.  Indeed, Jagdeesh and Titman argue that the momentum that they find in stock price returns is consistent with delayed price reaction to firm-specific news \citep{Jagadeesh}

@@

With a high level of uncertainty, there is a certain scepticism about new information and a caution about changing an established opinion. The Psychologist Edwards describes his findings relative to the benchmark of Bayesian decision-making.  
\begin{quotation}
``An abundance of research has shown that human beings are conservative processors of fallible information.  Such experiments compare human behaviour to the outputs of Bayes's theorem, the formally optimal rule about how opinions (that is, probabilities) should be revised on the basis of new information.  It turns out that opinion change is very \emph{orderly}, and usually proportional to number calculated from Bayes's theorem  - but it is insufficient in amount.  A convenient first approximation to the data would say that it takes anywhere from two to five observations to do one observation's worth of work in inducing a subject to change his opinions."  
\end{quotation}
\citep[p. 359]{KahnemanUncertainty}
